---
title: "sequencing_depth"
author: "Aakanksha Singh"
date: "2024-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(DropletUtils)
library(Matrix)
library(SingleCellExperiment)
library(org.Hs.eg.db)
library(Seurat)
source("./functions.R")
```


```{r}
setClassUnion("ExpData", c("matrix", "SummarizedExperiment"))
```


```{r}
processing_sce_modified_for_sequencing_depth <- function(dataset,title) 
 {   
   # Step 1: Quality Control
   stats <- high.mito <- list()
   unfiltered<-dataset
   
   current <- dataset
   #is.mito <- grep("MT", rowData(dataset)$Symbol_TENx)
   is.mito <- grep("MT", rowData(dataset)$Symbol)
   stats <- perCellQCMetrics(dataset, subsets=list(Mito=is.mito))
   high.mito <- isOutlier(stats$subsets_Mito_percent, type="higher")
   #dataset <- dataset[,!high.mito]
   dataset=dataset[,!high.mito & !is.na(high.mito)]
   
   current <- unfiltered 
   colData(current) <- cbind(colData(current), stats)
   current$discard <- high.mito
   
   # Note: The code line below generated a plot highlighting the mito genes as outliers
   qcplots<- plotColData(current, x="sum", y="subsets_Mito_percent",
                         colour_by="discard") + scale_x_log10()
   print("Processing sce : Step 1/7 Done")
   
   # Step 2: Log Normalisation
   dataset <- logNormCounts(dataset)
   print("Processing sce : Step 2/7 Done")
   
   # Step 3: Variance Modelling
   #dec <- modelGeneVar(dataset)
   # hvgs <-getTopHVGs(dataset, prop=0.1)
   # 
   # curdec <- dec
   # curfit <- metadata(curdec)
   # curve(curfit$trend(x), col='dodgerblue', lwd=2)
   print("Processing sce : Step 3/7 Done")
   
   # Step 4: Dimensionality Reduction (PCA, tSNE, UMAP)
   
   set.seed(10000)
   dataset <- runPCA(dataset)
   
   set.seed(100000)
   dataset <- runTSNE(dataset)
   
   set.seed(1000000)
   dataset <- runUMAP(dataset)
   print("Processing sce : Step 4/7 Done")
   
   # Step 5: Clustering
   g <- buildSNNGraph(dataset, k=10)
   clust <- igraph::cluster_walktrap(g)$membership
   colLabels(dataset)  <- factor(clust)
   print("Processing sce : Step 5/7 Done")
  
   # Step 6: Cell type annotation
   ref<-celldex::MonacoImmuneData(ensembl=TRUE, cell.ont="all")
   predicted_celltype <- SingleR(dataset, ref=ref, labels=ref$label.main)
   
   colData(dataset)$predicted_celltype <- predicted_celltype$labels
   tsne <- scater::plotTSNE(dataset, colour_by="predicted_celltype") + ggtitle(title)
   print("Processing sce : Step 6/7 Done")
   
   # Step 7: Adding an addition column that annotes all T cells the same
   # Adding an additional column with consolidated cell types, 
   # removing rows with no Hugo Symbol,
   # Setting the Hugo symbol as rownames instead of ensmbl ids
   
   dataset$cell<-dataset$predicted_celltype
   colData(dataset)$cell <- ifelse(colData(dataset)$cell %in% c("CD8+ T cells", "CD4+ T cells"), "T cells", colData(dataset)$cell)
   rownames(dataset) <- rowData(dataset)$Symbol
   dataset<-dataset[!(is.na(row.names(dataset))),]
   print("Processing sce : Step 7/7 Done")
   print(table(dataset$cell))
   
   return(dataset)
   
 }
```

## PART 1: analysing data from cell ranger ##
# Reading the 10x counts from multiplexed samples (To be done only once)
# The data is processed and provided in the intermediate_data folder
```{r}
# samples=c("bone_marrow_20k_0.8","bone_marrow_20k_0.6","bone_marrow_20k_0.4","bone_marrow_20k_0.2","bone_marrow_20k_0.01","bone_marrow_20k_0.001","bone_marrow_20k_0.0002","bone_marrow_20k_0.00002")
# 
# 
# for (sample_id in samples)
# {
#  
#     base_path= paste0("~/mnt_AG_Imkeller/group_members/asingh/sequencing_depth_datasets/bone_marrow_20k/outputs/cellranger/",sample_id,"/outs/per_sample_outs/")
#   print(paste0("Processing started for sample ",sample_id))
#   
#   sample_paths <- c(paste0(base_path,gsub("\\.","_", sample_id),"/count/sample_filtered_feature_bc_matrix"))
# 
#   # reading the count matrices
#   count_matrices <- lapply(sample_paths, function(path) {
#     Read10X(data.dir = path)
#   })
# 
#   # Aggregating the count matrices
#   for (i in seq_along(count_matrices)) {
#   colnames(count_matrices[[i]]) <- paste0(colnames(count_matrices[[i]]), "_Donor", i)
#   }
# 
#   combined_matrix <- do.call(cbind, count_matrices)
# 
#   # Converting the dcg matrix to SingleCellExperiment Object
#   sce_object <- SingleCellExperiment(
#     assays = list(counts = combined_matrix))
# 
#   # The rownames in the object are sometimes gene names other times the ENSEMNL ids, homogenising the names
#   gene_ids <- rownames(sce_object)
# 
#   rowdata_df= gene_ids %>%
#           as.data.frame() %>%
#           rename(Gene = ".") %>%
#           mutate(ENSEMBL_ID=ifelse(str_sub(Gene,0,4)=="ENSG",Gene,""),
#                  Symbol=ifelse(str_sub(Gene,0,4)!="ENSG",Gene,""))
# 
#   rowdata_df$ensid = mapIds(org.Hs.eg.db,
#                     keys=rowdata_df$Symbol,
#                     column="ENSEMBL",
#                     keytype="SYMBOL",
#                     multiVals="first")
# 
#   rowdata_df$symbol_from_ensembl = mapIds(org.Hs.eg.db,
#                     keys=rowdata_df$ENSEMBL_ID,
#                     column="SYMBOL",
#                     keytype="ENSEMBL",
#                     multiVals="first")
# 
#   rowdata_df=rowdata_df %>%
#     mutate(Symbol=ifelse(Symbol=="",symbol_from_ensembl,Symbol),
#            ENSEMBL_ID=ifelse(ENSEMBL_ID=="",ensid,ENSEMBL_ID)) %>%
#     dplyr::select(-ensid,-symbol_from_ensembl) %>%
#     mutate(original_name=Gene) %>%
#     column_to_rownames(var = "Gene") %>%
#     DataFrame()
# 
#   rowData(sce_object)=rowdata_df
# 
#   # removing the features that do not have a ensemblid in the dataset
#   sce_object=sce_object[is.na(rowData(sce_object)$ENSEMBL_ID)==FALSE,]
#   rownames(sce_object)=rowData(sce_object)$ENSEMBL_ID
# 
#   # Processing the SingleCellExperiment Object and saving it
#   processed_sce<-processing_sce_modified_for_sequencing_depth(sce_object,sample_id)
#   saveRDS(processed_sce, file = paste0("../intermediate_data/processed_sces/sequencing_depth/",sample_id,"_cell_ranger.rds"))
#   print(paste0("Processing completed for sample ", sample_id))
# }
# 

```

# Changing the annotations generated by Single R, so I take the annotations generated with 100% sequencing depth as the ground truth and perform subsetting based on that

PBMC
```{r}
pbmc20k_1 <- readRDS("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_1_cell_ranger.rds")
cell_count_pbmc20k_1 <- as.data.frame(table(pbmc20k_1$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

# Generating the "ground truth" in this case
ground_truth_cell_annotations=colData(pbmc20k_1) %>% 
                              as.data.frame() %>% 
                              dplyr::select(cell) %>% 
                              rownames_to_column(var="cell_barcode")

# Function for putting the ground truth in all the sce with the differential sequencing depths

changing_cell_annotations <- function(sequencing_depth,ground_truth_cell_annotations)
{
  sce_test <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_",sequencing_depth,"_cell_ranger.rds"))

  updated_coldata=colData(sce_test) %>% 
                as.data.frame() %>% 
                rownames_to_column(var="cell_barcode") %>% 
                dplyr::rename("independent_cell_annotation"=cell) %>% 
                left_join(ground_truth_cell_annotations,by=c("cell_barcode"="cell_barcode")) %>% 
                column_to_rownames(var="cell_barcode") %>% 
                DataFrame()
  colData(sce_test)=updated_coldata
  sce_test=sce_test[,is.na(sce_test$cell)==FALSE]
  return(sce_test)
}


pbmc20k_0.00002 <- changing_cell_annotations("0.00002",ground_truth_cell_annotations)
cell_count_pbmc20k_0.00002 <- as.data.frame(table(pbmc20k_0.00002$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.0002 <- changing_cell_annotations("0.0002",ground_truth_cell_annotations)
cell_count_pbmc20k_0.0002 <- as.data.frame(table(pbmc20k_0.0002$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.001 <- changing_cell_annotations("0.001",ground_truth_cell_annotations)
cell_count_pbmc20k_0.001 <- as.data.frame(table(pbmc20k_0.001$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.01 <- changing_cell_annotations("0.01",ground_truth_cell_annotations)
cell_count_pbmc20k_0.01 <- as.data.frame(table(pbmc20k_0.01$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.2 <- changing_cell_annotations("0.2",ground_truth_cell_annotations)
cell_count_pbmc20k_0.2 <- as.data.frame(table(pbmc20k_0.2$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.4 <- changing_cell_annotations("0.4",ground_truth_cell_annotations)
cell_count_pbmc20k_0.4 <- as.data.frame(table(pbmc20k_0.4$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.6 <- changing_cell_annotations("0.6",ground_truth_cell_annotations)
cell_count_pbmc20k_0.6 <- as.data.frame(table(pbmc20k_0.6$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.8 <- changing_cell_annotations("0.8",ground_truth_cell_annotations)
cell_count_pbmc20k_0.8 <- as.data.frame(table(pbmc20k_0.8$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.00002


```

# bone_marrow
```{r}
bone_marrow_20k_1 <- readRDS("../intermediate_data/processed_sces/sequencing_depth/bone_marrow_20k_1_cell_ranger.rds")
cell_count_bone_marrow_20k_1 <- as.data.frame(table(bone_marrow_20k_1$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

# Generating the "ground truth" in this case
ground_truth_cell_annotations=colData(bone_marrow_20k_1) %>% 
                              as.data.frame() %>% 
                              dplyr::select(cell) %>% 
                              rownames_to_column(var="cell_barcode")

# Function for putting the ground truth in all the sce with the differential sequencing depths

changing_cell_annotations <- function(sequencing_depth,ground_truth_cell_annotations)
{
  sce_test <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/bone_marrow_20k_",sequencing_depth,"_cell_ranger.rds"))

  updated_coldata=colData(sce_test) %>% 
                as.data.frame() %>% 
                rownames_to_column(var="cell_barcode") %>% 
                dplyr::rename("independent_cell_annotation"=cell) %>% 
                left_join(ground_truth_cell_annotations,by=c("cell_barcode"="cell_barcode")) %>% 
                column_to_rownames(var="cell_barcode") %>% 
                DataFrame()
  colData(sce_test)=updated_coldata
  sce_test=sce_test[,is.na(sce_test$cell)==FALSE]
  return(sce_test)
}


bone_marrow_20k_0.00002 <- changing_cell_annotations("0.00002",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.00002 <- as.data.frame(table(bone_marrow_20k_0.00002$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.0002 <- changing_cell_annotations("0.0002",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.0002 <- as.data.frame(table(bone_marrow_20k_0.0002$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.001 <- changing_cell_annotations("0.001",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.001 <- as.data.frame(table(bone_marrow_20k_0.001$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.01 <- changing_cell_annotations("0.01",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.01 <- as.data.frame(table(bone_marrow_20k_0.01$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.2 <- changing_cell_annotations("0.2",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.2 <- as.data.frame(table(bone_marrow_20k_0.2$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.4 <- changing_cell_annotations("0.4",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.4 <- as.data.frame(table(bone_marrow_20k_0.4$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.6 <- changing_cell_annotations("0.6",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.6 <- as.data.frame(table(bone_marrow_20k_0.6$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.8 <- changing_cell_annotations("0.8",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.8 <- as.data.frame(table(bone_marrow_20k_0.8$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

```


```{r}
# The dataset with random cells , date=10022025
dataset_included= c("pbmc20k_0.00002","pbmc20k_0.0002","pbmc20k_0.001","pbmc20k_0.01","pbmc20k_0.2","pbmc20k_0.4","pbmc20k_0.6","pbmc20k_0.8","pbmc20k_1")
cell_types= c("B cells","Monocytes","NK cells","T cells") 
bootstrap_value= 30
date="21052025"
data_type="pbmc20k_sequencing_depth_cell_ranger"
```


```{r}
dataset_included= c("bone_marrow_20k_0.00002","bone_marrow_20k_0.0002","bone_marrow_20k_0.001","bone_marrow_20k_0.01","bone_marrow_20k_0.2","bone_marrow_20k_0.4","bone_marrow_20k_0.6","bone_marrow_20k_0.8","bone_marrow_20k_1")
cell_types= c("B cells","Monocytes","NK cells","T cells") 
bootstrap_value= 30
date="04062025"
data_type="bone_marrow_20k_sequencing_depth_cell_ranger"
```

# Generating Bootstrapped samples Copied from generating bootstrapped samples)
# In future modify the generating_bootstrapped_samples and put it in the function file. 

```{r}
generating_bootstraped_datasets_10x <- function(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)
{  for (bootstrap in 1:bootstrap_value) 
    {
        for (dataset in dataset_included) 
          {
            var_name= paste0("cell_count_", dataset)
            cell_count_table <- get(var_name)
            sce_rds <- get(dataset)
            cell_table <-data.frame("cell_type"= cell_types) %>% 
                        left_join(cell_count_table, by = c("cell_type" = "cell_type")) %>% rowwise() %>% 
                        mutate(inDeconv = as.integer(runif(1,min=10,max = Freq))) %>%  as.data.frame()
            cell_table <- cell_table %>% mutate(inDeconv_Frac = inDeconv/sum(inDeconv))
            for_the_loop<- generating_a_pseudobulk_sample(sce_rds, 
                                                           dataset,                              
                                                           cell_table %>% pull(cell_type), 
                                                           cell_table %>% pull(inDeconv), 
                                                           perform_TPM_normalisation=FALSE)
            
           # Adding bootstrap value to the pseudobulks generated in each run
            for_the_loop= for_the_loop %>%
                dplyr::rename_with(~ paste0(.x,"_boot",bootstrap,".Raw"), -HUGO)
 
        
          if(dataset==dataset_included[1] && bootstrap==1)
            {
              deconv_table= for_the_loop
              actual_cell_proportions= cell_table %>% 
                             dplyr::select(cell_type,inDeconv,inDeconv_Frac) %>% 
                             dplyr::rename(!!paste0(dataset,"_boot",bootstrap,".actual_values") := "inDeconv",
                                           !!paste0(dataset,"_boot",bootstrap,".actual_frac") := "inDeconv_Frac")
                                 
            } else
            {
              deconv_table=merge(deconv_table, for_the_loop, by.x="HUGO",by.y="HUGO", all=TRUE)
              actual_cell_proportions= left_join(actual_cell_proportions, 
                                                 cell_table %>% 
                                                 dplyr::select(cell_type,inDeconv,inDeconv_Frac) %>% 
                                                 dplyr::rename(!!paste0(dataset,"_boot",bootstrap,".actual_values") := "inDeconv",
                                                               !!paste0(dataset,"_boot",bootstrap,".actual_frac") := "inDeconv_Frac"),
                                                 by=c("cell_type"="cell_type"))
           }
        }
        print(paste("Bootstrap",bootstrap,"of",bootstrap_value,"completed"))
      }
  saveRDS(deconv_table,file = paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_pseudobulks_",date,".RData"))
  saveRDS(actual_cell_proportions,file = paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_cell_proportions_",date,".RData"))  
}
```

```{r}
# To be run with 4 cell types
generating_bootstraped_datasets_10x(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)
```

```{r}
#For Epic (for 3 cell types)
dataset_included= c("pbmc20k_0.00002","pbmc20k_0.0002","pbmc20k_0.001","pbmc20k_0.01","pbmc20k_0.2","pbmc20k_0.4","pbmc20k_0.6","pbmc20k_0.8","pbmc20k_1")
cell_types= c("B cells","NK cells","T cells") 
bootstrap_value= 30
date="13102025"
data_type="pbmc20k_sequencing_depth_cell_ranger_for_epic"
generating_bootstraped_datasets_10x(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)


dataset_included= c("bone_marrow_20k_0.00002","bone_marrow_20k_0.0002","bone_marrow_20k_0.001","bone_marrow_20k_0.01","bone_marrow_20k_0.2","bone_marrow_20k_0.4","bone_marrow_20k_0.6","bone_marrow_20k_0.8","bone_marrow_20k_1")
cell_types= c("B cells","NK cells","T cells") 
bootstrap_value= 30
date="13102025"
data_type="bone_marrow_20k_sequencing_depth_cell_ranger_for_epic"

generating_bootstraped_datasets_10x(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)
```

# Analysis
PBMC
```{r}
date="13102025"
data_type="pbmc20k_sequencing_depth_cell_ranger"

# Pseudobulks generated from Chromium 10X
bootstraped_data=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_pseudobulks_21052025.RData"))
bootstraped_data_cell_counts=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_cell_proportions_21052025.RData")) %>%
                             dplyr::rename_with(~str_replace(., "_boot$", ""), ends_with("_boot"))


bootstraped_data_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_for_epic_pseudobulks_13102025.RData"))
bootstraped_data_cell_counts_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_for_epic_cell_proportions_13102025.RData")) %>%
                             dplyr::rename_with(~str_replace(., "_boot$", ""), ends_with("_boot"))

# Signature Matrix
signature_reference <- readRDS("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_1_cell_ranger.rds")
signature_reference$Barcode=colnames(signature_reference)

analysis_type="PBMC"
```

Bone Marrow
```{r}
data_type="bone_marrow_20k_sequencing_depth_cell_ranger"

# Pseudobulks generated from Chromium 10X
bootstraped_data=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_pseudobulks_04062025.RData"))
bootstraped_data_cell_counts=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_cell_proportions_04062025.RData")) %>%
                             dplyr::rename_with(~str_replace(., "_boot$", ""), ends_with("_boot"))
bootstraped_data_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_for_epic_pseudobulks_13102025.RData"))
bootstraped_data_cell_counts_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_for_epic_cell_proportions_13102025.RData")) %>%
                             dplyr::rename_with(~str_replace(., "_boot$", ""), ends_with("_boot"))

# Signature Matrix
signature_reference <- readRDS("../intermediate_data/processed_sces/sequencing_depth/bone_marrow_20k_1_cell_ranger.rds")
signature_reference$Barcode=colnames(signature_reference)

analysis_type="bone_marrow"
```


```{r}
date= "13102025"
column_character_specifications_for_mcp=c("0.00002","0.0002","0.001","0.01","0.2","0.4","0.6","0.8","_1_")
deconv_table=bootstraped_data %>%
                 mutate(across(-1, ~ (.x / sum(.x, na.rm = TRUE)) * 1e6))

deconv_table_epic=bootstraped_data_epic %>%
                 mutate(across(-1, ~ (.x / sum(.x, na.rm = TRUE)) * 1e6))

actual_cell_proportions= bootstraped_data_cell_counts
actual_cell_proportions_epic= bootstraped_data_cell_counts_epic
    
# Deconvolution
result_df <- deconvolution_for_bootstrap_v4(deconv_table, 
                                            deconv_tabe_epic,
                                            signature_reference,
                                            actual_cell_proportions,
                                            actual_cell_proportions_epic,
                                            column_character_specifications_for_mcp)

# Wrangling_data
comp_sequencing_depth_res= result_df %>%
      pivot_longer(cols = -c(deconv_type, cell_type),
                  names_to = c("Dataset", ".value"),
                  names_pattern = "^(.*_boot\\d+)\\.(Raw|actual_values|actual_frac)$") %>% 
                  # for real data
                  # names_sep = "\\.") %>% 
     dplyr::rename("Deconvolution_fraction"="Raw") %>% 
     dplyr::filter(!deconv_type %in% c("SpatialDecon (Custom)\n cells known","SpatialDecon (TME)\n cells known","SpatialDecon (Custom)\n cells unknown")) %>% 
                    mutate(deconv_type = ifelse(deconv_type=="SpatialDecon (TME)\n cells unknown","SpatialDecon",deconv_type)) %>% 
                    mutate(deconv_type=factor(deconv_type, levels=c("quanTIseq",
                                                                  "EPIC",
                                                                  "MCPcounter*",
                                                                  "SPOTlight",
                                                                  "SpatialDecon"))) %>% 
    mutate(Sequencing_depth= case_when(str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.00002"~"0.002% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.0002"~"0.02% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.001"~"0.1% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.01"~"1% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.2"~"20% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.4"~"40% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.6"~"60% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.8"~"80% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="1"~"100% reads")) %>%
    mutate(Sequencing_depth=factor(Sequencing_depth, levels=c("0.002% reads","0.02% reads","0.1% reads","1% reads","20% reads","40% reads","60% reads","80% reads","100% reads")))


save(comp_sequencing_depth_res, file = paste0("../Results/Sequencing_depth/cell_ranger/",analysis_type,"_sequencing_depth_",date,".RData"))
```

```{r}
load("../Results/Sequencing_depth/cell_ranger/PBMC_sequencing_depth_pbmc20k_21052025.RData") 
```

# Checking the depth of the pseudo-bulk sample
```{r}
sequencing_depth_of_pseudo_bulks <- tibble(
  "Sequencing_depth_prop" = rev(c("0.00002", "0.0002", "0.01", "0.2", "0.6", "_1_")),
  "Sequencing depth \n of pseudo-bulk samples" = rev(c("0.002% reads", "0.02% reads", "1% reads", "20% reads", "60% reads", "100% reads"))
)

bootstraped_data_pbmc=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/pbmc20k_sequencing_depth_cell_ranger_pseudobulks_21052025.RData"))
bootstraped_data_pbmc_cell=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/pbmc20k_sequencing_depth_cell_ranger_cell_proportions_21052025.RData"))


bootstraped_data_pbmc_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/pbmc20k_sequencing_depth_cell_ranger_for_epic_pseudobulks_13102025.RData"))
bootstraped_data_pbmc_cell_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/pbmc20k_sequencing_depth_cell_ranger_for_epic_cell_proportions_13102025.RData"))


bootstraped_data_bone_marrow=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/bone_marrow_20k_sequencing_depth_cell_ranger_pseudobulks_04062025.RData"))
bootstraped_data_bone_marrow_cells=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/bone_marrow_20k_sequencing_depth_cell_ranger_cell_proportions_04062025.RData"))

bootstraped_data_bone_marrow_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/bone_marrow_20k_sequencing_depth_cell_ranger_for_epic_pseudobulks_13102025.RData"))
bootstraped_data_bone_marrow_cells_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/bone_marrow_20k_sequencing_depth_cell_ranger_for_epic_cell_proportions_13102025.RData"))

#i="0.00002"


for (i in sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop) 
{
  # PBMC 4 cells
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean counts"] <-
    bootstraped_data_pbmc %>%
      select(contains(i)) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean nb. of cells"] <-
    bootstraped_data_pbmc_cell %>%
      select(contains(i)) %>%
      select(contains("actual_values")) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean nb. of genes"] <-
    bootstraped_data_pbmc %>%
      select(contains(i)) %>%
      summarise(across(everything(), ~ sum(. != 0))) %>%  # count non-zero per column
      unlist() %>%                                       # convert to numeric vector
      mean() 
  
  # PBMC 3 cells
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean counts \n3 cells in pseudobulks"] <-
    bootstraped_data_pbmc_for_epic %>%
      select(contains(i)) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean nb. of cells \n3 cells in pseudobulks"] <-
    bootstraped_data_pbmc_cell_for_epic %>%
      select(contains(i)) %>%
      select(contains("actual_values")) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean nb. of genes \n3 cells in pseudobulks"] <-
    bootstraped_data_pbmc_for_epic %>%
      select(contains(i)) %>%
      summarise(across(everything(), ~ sum(. != 0))) %>%  # count non-zero per column
      unlist() %>%                                       # convert to numeric vector
      mean() 
  
  # Bone Marrow 4 cells
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean counts"] <-
    bootstraped_data_bone_marrow %>%
      select(contains(i)) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean nb. of cells"] <-
    bootstraped_data_bone_marrow_cells %>%
      select(contains(i)) %>%
      select(contains("actual_values")) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean nb. of genes"] <-
    bootstraped_data_bone_marrow %>%
      select(contains(i)) %>%
      summarise(across(everything(), ~ sum(. != 0))) %>%  # count non-zero per column
      unlist() %>%                                       # convert to numeric vector
      mean() 
  
  # Bone Marrow 3 cells
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean counts \n 3cells in pseudobulks"] <-
    bootstraped_data_bone_marrow_for_epic %>%
      select(contains(i)) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean nb. of cells \n 3cells in pseudobulks"] <-
    bootstraped_data_bone_marrow_cells_for_epic %>%
      select(contains(i)) %>%
      select(contains("actual_values")) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean nb. of genes \n 3cells in pseudobulks"] <-
    bootstraped_data_bone_marrow_for_epic %>%
      select(contains(i)) %>%
      summarise(across(everything(), ~ sum(. != 0))) %>%  # count non-zero per column
      unlist() %>%                                       # convert to numeric vector
      mean() 
}

sequencing_depth_of_pseudo_bulks=sequencing_depth_of_pseudo_bulks %>% 
  select(-Sequencing_depth_prop)

saveRDS(sequencing_depth_of_pseudo_bulks,"../manuscript/figure_assembly_scripts/fig_sequencing_depth/sequencing_depth_summary_numbers.RData")

library(ggplot2)
library(gridExtra)

png("../manuscript/figure_assembly_scripts/fig_sequencing_depth/sequencing_depth_summary_numbers.png", height=180, width=800,pointsize = 16)
p<-tableGrob(sequencing_depth_of_pseudo_bulks)
grid.arrange(p)
dev.off()
```
