---
title: "Step1: Generating_dataset_with_pseudobulks"
author: "Aakanksha Singh"
date: "2024-05-13"
output: html_document
---

This is the first step of the Deconvolution Benchmarking pipeline. In this script 
our aim is to 
1. get the single cell data from the publicly available repositories and process it.
2. Generate a countmatrix with bootstrapped single cell data.
3. Generate 4 of such count matrices
  a. PBMC from 10X chromium (For Gene Panel, Normalisation,Signature Matrix and Count matrix origin)
  b. PBMC data from SmartSeq2 (For count matrix origin)
  c. Test dataset of another kind (10X Chromium) AML
  d. Test dataset of another kind (Smart Seq2) AML
  
Requirements for generating the pseudobulk samples: Each dataset for each type should have a minimum of 10 cells in the
cell types that are being examined in the analysis ie., ( B cells, T cells, Monocytes and NK cells)

```{r}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install(version = "3.19")
# BiocManager::install("TENxPBMCData")
# BiocManager::install("scater")
# BiocManager::install("scran")
# BiocManager::install("celldex")
# update.packages(checkBuilt = TRUE, ask = FALSE)
source("./functions.R")
```

# Dataset1: 10X Chromium PBMC data

## Reading data, Annotating cell types and saving it (Only to be done once)
```{r}
# pbmc3k<-processing_sce(TENxPBMCData("pbmc3k"),"pbmc3k")
# saveRDS(pbmc3k, file = "../intermediate_data/processed_sces/pbmc3k.rds")
# 
# pbmc4k<-processing_sce(TENxPBMCData("pbmc4k"),"pbmc4k")
# saveRDS(pbmc4k, file = "../intermediate_data/processed_sces/pbmc4k.rds")
#
# pbmc8k<-processing_sce(TENxPBMCData("pbmc8k"),"pbmc8k")
# saveRDS(pbmc8k, file = "../intermediate_data/processed_sces/pbmc8k.rds")
# 
# pbmc6k<-processing_sce(TENxPBMCData("pbmc6k"),"pbmc6k")
# saveRDS(pbmc6k, file = "../intermediate_data/processed_sces/pbmc6k.rds")
# 
# liver<-processing_sce(ZhaoImmuneLiverData(),"liver")
# saveRDS(liver, file = "../intermediate_data/processed_sces/liver.rds")
```

## Loading the Data and calculating the number of cells in each cell type
```{r}
pbmc3k <- readRDS("../intermediate_data/processed_sces/pbmc3k.rds")
cell_count_pbmc3k <- as.data.frame(table(pbmc3k$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc4k <- readRDS("../intermediate_data/processed_sces/pbmc4k.rds")
cell_count_pbmc4k <- as.data.frame(table(pbmc4k$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc6k <- readRDS("../intermediate_data/processed_sces/pbmc6k.rds")
cell_count_pbmc6k <- as.data.frame(table(pbmc6k$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc8k <- readRDS("../intermediate_data/processed_sces/pbmc8k.rds")
cell_count_pbmc8k <- as.data.frame(table(pbmc8k$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")


# genes_pbmc8k =rownames(pbmc8k) %>% 
#               as.data.frame()
# write.csv(genes_pbmc8k, file = "../Data/Signature_matrices/custom_signature_pbmc8k_genes.csv", col.names = FALSE)
# write.csv(safeTME %>% as.data.frame(),file = "../Data/Signature_matrices/safeTME.csv", col.names = FALSE)
# lm22=read.delim("../Data/Signature_matrices/LM22_signature_matrix.txt")
# write.csv(lm22,file ="../Data/Signature_matrices/LM22_signature_matrix.csv" )
```

```{r}
dataset_included= c("pbmc3k","pbmc4k","pbmc6k","pbmc8k")
cell_types= c("B cells","Monocytes","NK cells","T cells")
bootstrap_value= 30
#date="03072024"
date=24022025
data_type="pbmc_10x"
```

```{r}
generating_bootstraped_datasets_10x <- function(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)
{  for (bootstrap in 1:bootstrap_value) 
    {
        for (dataset in dataset_included) 
          {
            var_name= paste0("cell_count_", dataset)
            cell_count_table <- get(var_name)
            sce_rds <- get(dataset)
            cell_table <-data.frame("cell_type"= cell_types) %>% 
                        left_join(cell_count_table, by = c("cell_type" = "cell_type")) %>% rowwise() %>% 
                        mutate(inDeconv = as.integer(runif(1,min=10,max = Freq))) %>%  as.data.frame()
            cell_table <- cell_table %>% mutate(inDeconv_Frac = inDeconv/sum(inDeconv))
            for_the_loop<- generating_a_pseudobulk_sample(
                                                           #sce_rds[rowData(sce_rds)$Symbol], 
                                                           sce_rds,
                                                           dataset,                              
                                                           cell_table %>% pull(cell_type), 
                                                           cell_table %>% pull(inDeconv), 
                                                           perform_TPM_normalisation=FALSE)
            
           # Adding bootstrap value to the pseudobulks generated in each run
            for_the_loop= for_the_loop %>%
                dplyr::rename_with(~ paste0(.x,"_boot",bootstrap,".Raw"), -HUGO)
 
        
          if(dataset==dataset_included[1] && bootstrap==1)
            {
              deconv_table= for_the_loop
              actual_cell_proportions= cell_table %>% 
                             dplyr::select(cell_type,inDeconv,inDeconv_Frac) %>% 
                             dplyr::rename(!!paste0(dataset,"_boot",bootstrap,".actual_values") := "inDeconv",
                                           !!paste0(dataset,"_boot",bootstrap,".actual_frac") := "inDeconv_Frac")
                                 
            } else
            {
              deconv_table=merge(deconv_table, for_the_loop, by.x="HUGO",by.y="HUGO", all=TRUE)
              actual_cell_proportions= left_join(actual_cell_proportions, 
                                                 cell_table %>% 
                                                 dplyr::select(cell_type,inDeconv,inDeconv_Frac) %>% 
                                                 dplyr::rename(!!paste0(dataset,"_boot",bootstrap,".actual_values") := "inDeconv",
                                                               !!paste0(dataset,"_boot",bootstrap,".actual_frac") := "inDeconv_Frac"),
                                                 by=c("cell_type"="cell_type"))
           }
        }
        print(paste("Bootstrap",bootstrap,"of",bootstrap_value,"completed"))
      }
  saveRDS(deconv_table,file = paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_pseudobulks_",date,".RData"))
  saveRDS(actual_cell_proportions,file = paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_cell_proportions_",date,".RData"))  
}
```

#### For 4 cell types#####
```{r}
generating_bootstraped_datasets_10x(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)
```


#### For 3 cell types ####
```{r}

dataset_included= c("pbmc3k","pbmc4k","pbmc6k","pbmc8k")
bootstrap_value= 30
date=13102025
data_type="pbmc_10x_for_epic"

generating_bootstraped_datasets_10x(dataset_included,
                                            data_type,  # i.e with 3 cell_types
                                            c("B cells","NK cells","T cells"),
                                            bootstrap_value,
                                            date)
```

# With a max cell limit
Generating a pseudobulk sample to see if the number of cells in a sample impacts the deconvolution accuracy
```{r}
dataset_included= c("pbmc4k","pbmc6k","pbmc8k")
#cell_types= c("B cells","Monocytes","NK cells","T cells")
cell_types=c("B cells","NK cells","T cells")
bootstrap_value= 30
#date="01102025"
date="13102025"
data_type="pbmc_for_epic"

# max_cells=c(10,50,150)
# min_cells=c(2,10,50)

# per cell
max_cells=c(4,125,1000)
min_cells=c(2,25,200)

# bootstrap=1
# sample_size=2
# dataset="pbmc6k"


# generating_bootstraped_datasets_10x <- function(dataset_included,
#                                             data_type,
#                                             cell_types,
#                                             bootstrap_value,
#                                             date)


  for (bootstrap in 1:bootstrap_value) 
    { 
      for (sample_size in 1:3)
       {
        for (dataset in dataset_included) 
          {
            var_name= paste0("cell_count_", dataset)
            cell_count_table <- get(var_name)
            sce_rds <- get(dataset)
            cell_table <-data.frame("cell_type"= cell_types) %>% 
                        left_join(cell_count_table, by = c("cell_type" = "cell_type")) %>% rowwise() %>% 
              mutate(inDeconv = sample(min_cells[sample_size]:ifelse(max(.data$Freq) < max_cells[sample_size], max(.data$Freq), max_cells[sample_size]), size = n(), replace = FALSE)) %>% as.data.frame()

            cell_table <- cell_table %>% mutate(inDeconv_Frac = inDeconv/sum(inDeconv))
            for_the_loop<- generating_a_pseudobulk_sample(sce_rds[rowData(sce_rds)$Symbol], 
                                                           paste0(dataset,".",max_cells[sample_size]),                              
                                                           cell_table %>% pull(cell_type), 
                                                           cell_table %>% pull(inDeconv), 
                                                           perform_TPM_normalisation=FALSE)
            
           # Adding bootstrap value to the pseudobulks generated in each run
            for_the_loop= for_the_loop %>%
                dplyr::rename_with(~ paste0(.x,"_boot",bootstrap,".Raw"), -HUGO)
 
        
          if(dataset==dataset_included[1] && bootstrap==1)
            {
              deconv_table= for_the_loop
              actual_cell_proportions= cell_table %>% 
                             dplyr::select(cell_type,inDeconv,inDeconv_Frac) %>% 
                             dplyr::rename(!!paste0(dataset,".",max_cells[sample_size],"_boot",bootstrap,".actual_values") := "inDeconv",
                                           !!paste0(dataset,".",max_cells[sample_size],"_boot",bootstrap,".actual_frac") := "inDeconv_Frac")
                                 
            } else
            {
              deconv_table=merge(deconv_table, for_the_loop, by.x="HUGO",by.y="HUGO", all=TRUE)
              actual_cell_proportions= left_join(actual_cell_proportions, 
                                                 cell_table %>% 
                                                 dplyr::select(cell_type,inDeconv,inDeconv_Frac) %>% 
                                                 dplyr::rename(!!paste0(dataset,".",max_cells[sample_size],"_boot",bootstrap,".actual_values") := "inDeconv",
                                                               !!paste0(dataset,".",max_cells[sample_size],"_boot",bootstrap,".actual_frac") := "inDeconv_Frac"),
                                                 by=c("cell_type"="cell_type"))
           }
        }
        print(paste("Bootstrap",bootstrap,"of",bootstrap_value,"and Sample size",sample_size,"of",length(max_cells),"completed"))
      }
  }
  
  saveRDS(deconv_table,file = paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_pseudobulks_spot_size_",date,".RData"))
  saveRDS(actual_cell_proportions,file = paste0("../intermediate_data/bootstraped_pseudobulks/",data_type,"_cell_proportions_spot_size_",date,".RData")) 
```


```{r}
# df=readRDS("../intermediate_data/bootstraped_pseudobulks/pbmc_pseudobulks_13052024.RData")
# colnames(df)
```

## AML Datasets to go in the Supplementary:
To make our results reproducible on a different dataset: we use the AML dataset

# Dataset3: 10X Chromium AML data
```{r}
AML0110X <- readRDS("../intermediate_data/processed_sces/AML0110X.rds")
cell_count_AML0110X <- as.data.frame(table(AML0110X$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

AML0210X <- readRDS("../intermediate_data/processed_sces/AML0210X.rds")
cell_count_AML0210X <- as.data.frame(table(AML0210X$cell)) %>% 
                       dplyr::rename("cell_type"="Var1")

AML0310X <- readRDS("../intermediate_data/processed_sces/AML0310X.rds")
cell_count_AML0310X <- as.data.frame(table(AML0310X$cell)) %>% 
                       dplyr::rename("cell_type"="Var1")

signature_reference_aml_10X= readRDS("../intermediate_data/processed_sces/AML0110X.rds")
```


```{r}
dataset_included= c("AML0110X","AML0210X","AML0310X")
cell_types= c("B cells","Monocytes","NK cells","T cells")
bootstrap_value= 30
date="03072024"
data_type="aml_10x"

generating_bootstraped_datasets_10x(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)
```

```{r}
dataset_included= c("AML0110X","AML0210X","AML0310X")
cell_types= c("B cells","NK cells","T cells") # for epic
bootstrap_value= 30
date="13102025"
data_type="aml_10x_for_epic"

generating_bootstraped_datasets_10x(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)
```
