---
title: "sequencing_depth"
author: "Aakanksha Singh"
date: "2024-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(DropletUtils)
library(Matrix)
library(SingleCellExperiment)
library(org.Hs.eg.db)
library(Seurat)
source("./0_functions.R")
```

```{r}
setClassUnion("ExpData", c("matrix", "SummarizedExperiment"))
```

```{r}
#dataset = sce_object

processing_sce_modified_for_sequencing_depth <- function(dataset,title) 
 {   
   
   # # Step 1 and 2 to be commented out when using down sampled instances
   # # Log Normalisation and Cell type annotation

   # dataset <- logNormCounts(dataset)
   # print("Processing sce : Step 2/7 Done")
   # ref<-celldex::MonacoImmuneData(ensembl=TRUE, cell.ont="all")
   # predicted_celltype <- SingleR(dataset, ref=ref, labels=ref$label.main)
   # colData(dataset)$predicted_celltype <- predicted_celltype$labels
   # 
   # # Step 4: Adding an addition column that annotes all T cells the same
   # # Adding an additional column with consolidated cell types,
   # # removing rows with no Hugo Symbol,
   # # Setting the Hugo symbol as rownames instead of ensmbl ids
   # 
   # dataset$cell<-dataset$predicted_celltype
   # colData(dataset)$cell <- ifelse(colData(dataset)$cell %in% c("CD8+ T cells", "CD4+ T cells"), "T cells", colData(dataset)$cell)
   rownames(dataset) <- rowData(dataset)$Symbol
   dataset<-dataset[!(is.na(row.names(dataset))),]
   return(dataset)
   
 }
```

## PART 1: analysing data from cell ranger ##
# Reading the 10x counts from multiplexed samples (To be done only once)
# The data is processed and provided in the intermediate_data folder
```{r}
# Bone Marrow
tissue_type="bone_marrow_20k"
#sample_id="bone_marrow_20k_1"
samples=c("bone_marrow_20k_0.8","bone_marrow_20k_0.6","bone_marrow_20k_0.4","bone_marrow_20k_0.2","bone_marrow_20k_0.01","bone_marrow_20k_0.001","bone_marrow_20k_0.0002","bone_marrow_20k_0.00002")

# tissue_type="PBMC20k"
# sample_id="pbmc20k_1"
#samples=c("pbmc20k_0.8","pbmc20k_0.6","pbmc20k_0.4","pbmc20k_0.2","pbmc20k_0.01","pbmc20k_0.001","pbmc20k_0.0002","pbmc20k_0.00002")


for (sample_id in samples)
{

  base_path <- paste0("../Data/sequencing_depth_datasets/", tissue_type, "/Output_data/cell_ranger/", sample_id, "/outs/per_sample_outs/")
print(paste0("Processing started for sample ", sample_id))

if (sample_id == "pbmc20k_1" & tissue_type == "PBMC20k") {
  sample_paths <- c(paste0(base_path, "5k_Human_Donor1_PBMC_3p_gem-x/count/sample_filtered_feature_bc_matrix"),
                    paste0(base_path, "5k_Human_Donor2_PBMC_3p_gem-x/count/sample_filtered_feature_bc_matrix"),
                    paste0(base_path, "5k_Human_Donor3_PBMC_3p_gem-x/count/sample_filtered_feature_bc_matrix"),
                    paste0(base_path, "5k_Human_Donor4_PBMC_3p_gem-x/count/sample_filtered_feature_bc_matrix"))

} else if (tissue_type == "PBMC20k" & sample_id != "pbmc20k_1") {
  sample_paths <- c(paste0(base_path, "5k_Human_Donor1_PBMC_3p_gem-x/count/sample_raw_feature_bc_matrix"),
                    paste0(base_path, "5k_Human_Donor2_PBMC_3p_gem-x/count/sample_raw_feature_bc_matrix"),
                    paste0(base_path, "5k_Human_Donor3_PBMC_3p_gem-x/count/sample_raw_feature_bc_matrix"),
                    paste0(base_path, "5k_Human_Donor4_PBMC_3p_gem-x/count/sample_raw_feature_bc_matrix"))

} else if (sample_id == "bone_marrow_20k_1" & tissue_type == "bone_marrow_20k") {
  sample_paths <- c(paste0(base_path, gsub("\\.", "_", sample_id), "/count/sample_filtered_feature_bc_matrix"))

} else if (tissue_type == "bone_marrow_20k" & sample_id != "bone_marrow_20k_1") {
  sample_paths <- c(paste0("../Data/sequencing_depth_datasets/", tissue_type, "/Output_data/cell_ranger/", sample_id, "/outs/multi/count/raw_feature_bc_matrix"))
}
  
  
  #sample_paths <- c(paste0(base_path,gsub("\\.","_", sample_),"/count/sample_raw_feature_bc_matrix"))

  #sample_paths <- c(paste0(base_path,gsub("\\.","_", sample_id),"/count/sample_raw_feature_bc_matrix"))

  # reading the count matrices
  count_matrices <- lapply(sample_paths, function(path) {
    Read10X(data.dir = path)
  })

  # Aggregating the count matrices
  for (i in seq_along(count_matrices)) {
  colnames(count_matrices[[i]]) <- paste0(colnames(count_matrices[[i]]), "_Donor", i)
  }

  combined_matrix <- do.call(cbind, count_matrices)

  # Converting the dcg matrix to SingleCellExperiment Object
  sce_object <- SingleCellExperiment(
    assays = list(counts = combined_matrix))

  # The rownames in the object are sometimes gene names other times the ENSEMNL ids, homogenising the names
  gene_ids <- rownames(sce_object)

  rowdata_df= gene_ids %>%
          as.data.frame() %>%
          rename(Gene = ".") %>%
          mutate(ENSEMBL_ID=ifelse(str_sub(Gene,0,4)=="ENSG",Gene,""),
                 Symbol=ifelse(str_sub(Gene,0,4)!="ENSG",Gene,""))

  rowdata_df$ensid = mapIds(org.Hs.eg.db,
                    keys=rowdata_df$Symbol,
                    column="ENSEMBL",
                    keytype="SYMBOL",
                    multiVals="first")

  rowdata_df$symbol_from_ensembl = mapIds(org.Hs.eg.db,
                    keys=rowdata_df$ENSEMBL_ID,
                    column="SYMBOL",
                    keytype="ENSEMBL",
                    multiVals="first")

  rowdata_df=rowdata_df %>%
    mutate(Symbol=ifelse(Symbol=="",symbol_from_ensembl,Symbol),
           ENSEMBL_ID=ifelse(ENSEMBL_ID=="",ensid,ENSEMBL_ID)) %>%
    dplyr::select(-ensid,-symbol_from_ensembl) %>%
    mutate(original_name=Gene) %>%
    column_to_rownames(var = "Gene") %>%
    DataFrame()

  rowData(sce_object)=rowdata_df

  # removing the features that do not have a ensemblid in the dataset
  sce_object=sce_object[is.na(rowData(sce_object)$ENSEMBL_ID)==FALSE,]
  rownames(sce_object)=rowData(sce_object)$ENSEMBL_ID

  # Processing the SingleCellExperiment Object and saving it
  processed_sce<-processing_sce_modified_for_sequencing_depth(sce_object,sample_id)
  saveRDS(processed_sce, file = paste0("../intermediate_data/processed_sces/sequencing_depth/",sample_id,"_cell_ranger.rds"))
  print(paste0("Processing completed for sample ", sample_id))
}
```

# Changing the annotations generated by Single R, so I take the annotations generated with 100% sequencing depth as the ground truth and perform subsetting based on that

PBMC
```{r}
pbmc20k_1 <- readRDS("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_1_cell_ranger.rds")

cell_count_pbmc20k_1 <- as.data.frame(table(pbmc20k_1$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

# Generating the "ground truth" in this case
ground_truth_cell_annotations=colData(pbmc20k_1) %>% 
                              as.data.frame() %>% 
                              dplyr::select(cell) %>% 
                              rownames_to_column(var="cell_barcode")

# Function for putting the ground truth in all the sce with the differential sequencing depths
adding_cell_annotations <- function(sequencing_depth,ground_truth_cell_annotations)
{
  sce_test <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_",sequencing_depth,"_cell_ranger.rds"))

  updated_coldata=colData(sce_test) %>% 
                as.data.frame() %>% 
                rownames_to_column(var="cell_barcode") %>% 
                #dplyr::rename("independent_cell_annotation"=cell) %>% 
                left_join(ground_truth_cell_annotations,by=c("cell_barcode"="cell_barcode")) %>% 
                column_to_rownames(var="cell_barcode") %>% 
                DataFrame()
  colData(sce_test)=updated_coldata
  sce_test=sce_test[,is.na(sce_test$cell)==FALSE]
  return(sce_test)
}


pbmc20k_0.00002 <- adding_cell_annotations("0.00002",ground_truth_cell_annotations)
cell_count_pbmc20k_0.00002 <- as.data.frame(table(pbmc20k_0.00002$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.0002 <- adding_cell_annotations("0.0002",ground_truth_cell_annotations)
cell_count_pbmc20k_0.0002 <- as.data.frame(table(pbmc20k_0.0002$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.001 <- adding_cell_annotations("0.001",ground_truth_cell_annotations)
cell_count_pbmc20k_0.001 <- as.data.frame(table(pbmc20k_0.001$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.01 <- adding_cell_annotations("0.01",ground_truth_cell_annotations)
cell_count_pbmc20k_0.01 <- as.data.frame(table(pbmc20k_0.01$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.2 <- adding_cell_annotations("0.2",ground_truth_cell_annotations)
cell_count_pbmc20k_0.2 <- as.data.frame(table(pbmc20k_0.2$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.4 <- adding_cell_annotations("0.4",ground_truth_cell_annotations)
cell_count_pbmc20k_0.4 <- as.data.frame(table(pbmc20k_0.4$cell)) %>%
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.6 <- adding_cell_annotations("0.6",ground_truth_cell_annotations)
cell_count_pbmc20k_0.6 <- as.data.frame(table(pbmc20k_0.6$cell)) %>%
                     dplyr::rename("cell_type"="Var1")

pbmc20k_0.8 <- adding_cell_annotations("0.8",ground_truth_cell_annotations)
cell_count_pbmc20k_0.8 <- as.data.frame(table(pbmc20k_0.8$cell)) %>%
                     dplyr::rename("cell_type"="Var1")
```

```{r}
sce_list=list("pbmc_1"=pbmc20k_1,"pbmc_0.8"=pbmc20k_0.8,"pbmc_0.6"=pbmc20k_0.6,"pbmc_0.4"=pbmc20k_0.4,"pbmc_0.2"=pbmc20k_0.2,"pbmc_0.01"=pbmc20k_0.01,"pbmc_0.001"=pbmc20k_0.001,"pbmc_0.0002"=pbmc20k_0.0002,"pbmc_0.00002"=pbmc20k_0.00002)

barcode_counts <- sapply(sce_list, function(sce) sum(colnames(pbmc20k_1) %in% colnames(sce)))

sc_barcodes_in_subsamples <- data.frame(
    Sample = names(sce_list),
    Matching_barcodes = barcode_counts
)
```

```{r}
# sce_1 <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_1_cell_ranger.rds"))
# sce_0.8 <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_0.8_cell_ranger.rds"))
# sce_0.6 <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_0.6_cell_ranger.rds"))
# sce_0.4 <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_0.4_cell_ranger.rds"))
# sce_0.2 <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_0.2_cell_ranger.rds"))
# sce_0.01 <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_0.01_cell_ranger.rds"))
# sce_0.001 <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_0.001_cell_ranger.rds"))
# sce_0.0002 <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_0.0002_cell_ranger.rds"))
# sce_0.00002 <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_0.00002_cell_ranger.rds"))
# 
# real_cells_df<-data_frame()
# real_cell_numbers <-function("sample_id") 
#   {
#   unfiltered <- get(paste0("pbmc20k_",sample_id))
#   filtered <- readRDS(paste0("~/mnt_AG_Imkeller/group_members/asingh/Benchmarking_deconvolution_update/Older_version/Benchmarking_deconvolution_v1/intermediate_data/processed_sces/sequencing_depth/PBMC20k_",sample_id,"_cell_ranger.rds"))
#   real_cells_df[["Sample"]]=paste0("pbmc20k_",sample_id)
#   real_cells_df[["Nb. of Unfiltered barcodes"]]=ncol(unfiltered)
#   real_cells_df[["Nb. of Filtered barcodes"]]=ncol(filtered)
#   real_cells_df[["Nb. of real cells in unfiltered"]]= ncol(unfiltered[,colnames(filtered)])
# }
# 
# rm(real_cell_numbers_df)
# 
# filtered <- readRDS(paste0("~/mnt_AG_Imkeller/group_members/asingh/Benchmarking_deconvolution_update/Older_version/Benchmarking_deconvolution_v1/intermediate_data/processed_sces/sequencing_depth/PBMC20k_1_cell_ranger.rds"))
# 
# real_cell_numbers <- function(sample_id) {
#   unfiltered <- get(paste0("pbmc20k_", sample_id))
#   
#   real_cells_df <- list()
#   real_cells_df[["Sample"]] <- paste0("pbmc20k_", sample_id)
#   real_cells_df[["Nb. of Unfiltered barcodes"]] <- ncol(unfiltered)
#   real_cells_df[["Nb. of Filtered barcodes"]] <- ncol(filtered)
#   real_cells_df[["Nb. of real cells in unfiltered"]] <- ncol(unfiltered[, colnames(unfiltered) %in% colnames(filtered)])
#   
#   return(real_cells_df)
# }
# 
# real_cell_numbers_df <- do.call(rbind, lapply(c("1", "0.8", "0.6", "0.4","0.2","0.01","0.001","0.0002","0.00002"), function(sid) {
#   as.data.frame(real_cell_numbers(sid))
# }))
# 
# 
# colnames(assay(pbmc20k_1[1:5,1:5]))
```

# bone_marrow
```{r}
bone_marrow_20k_1 <- readRDS("../intermediate_data/processed_sces/sequencing_depth/bone_marrow_20k_1_cell_ranger.rds")
cell_count_bone_marrow_20k_1 <- as.data.frame(table(bone_marrow_20k_1$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

# Generating the "ground truth" in this case
ground_truth_cell_annotations=colData(bone_marrow_20k_1) %>% 
                              as.data.frame() %>% 
                              dplyr::select(cell) %>% 
                              rownames_to_column(var="cell_barcode")


# Function for putting the ground truth in all the sce with the differential sequencing depths
adding_cell_annotations <- function(sequencing_depth,ground_truth_cell_annotations)
{
  sce_test <- readRDS(paste0("../intermediate_data/processed_sces/sequencing_depth/bone_marrow_20k_",sequencing_depth,"_cell_ranger.rds"))

  updated_coldata=colData(sce_test) %>% 
                as.data.frame() %>% 
                rownames_to_column(var="cell_barcode") %>% 
                #dplyr::rename("independent_cell_annotation"=cell) %>% 
                left_join(ground_truth_cell_annotations,by=c("cell_barcode"="cell_barcode")) %>% 
                column_to_rownames(var="cell_barcode") %>% 
                DataFrame()
  colData(sce_test)=updated_coldata
  sce_test=sce_test[,is.na(sce_test$cell)==FALSE]
  return(sce_test)
}

bone_marrow_20k_0.00002 <- adding_cell_annotations("0.00002",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.00002 <- as.data.frame(table(bone_marrow_20k_0.00002$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.0002 <- adding_cell_annotations("0.0002",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.0002 <- as.data.frame(table(bone_marrow_20k_0.0002$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.001 <- adding_cell_annotations("0.001",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.001 <- as.data.frame(table(bone_marrow_20k_0.001$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.01 <- adding_cell_annotations("0.01",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.01 <- as.data.frame(table(bone_marrow_20k_0.01$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.2 <- adding_cell_annotations("0.2",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.2 <- as.data.frame(table(bone_marrow_20k_0.2$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.4 <- adding_cell_annotations("0.4",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.4 <- as.data.frame(table(bone_marrow_20k_0.4$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.6 <- adding_cell_annotations("0.6",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.6 <- as.data.frame(table(bone_marrow_20k_0.6$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

bone_marrow_20k_0.8 <- adding_cell_annotations("0.8",ground_truth_cell_annotations)
cell_count_bone_marrow_20k_0.8 <- as.data.frame(table(bone_marrow_20k_0.8$cell)) %>% 
                     dplyr::rename("cell_type"="Var1")

```


```{r}
# The dataset with random cells , date=10022025
dataset_included= c("pbmc20k_0.00002","pbmc20k_0.0002","pbmc20k_0.001","pbmc20k_0.01","pbmc20k_0.2","pbmc20k_0.4","pbmc20k_0.6","pbmc20k_0.8")
cell_types= c("B cells","Monocytes","NK cells","T cells") 
bootstrap_value= 30
date="09042026"
data_type="pbmc20k_sequencing_depth_cell_ranger"
```


```{r}
dataset_included= c("bone_marrow_20k_0.00002","bone_marrow_20k_0.0002","bone_marrow_20k_0.001","bone_marrow_20k_0.01","bone_marrow_20k_0.2","bone_marrow_20k_0.4","bone_marrow_20k_0.6","bone_marrow_20k_0.8")
cell_types= c("B cells","Monocytes","NK cells","T cells") 
bootstrap_value= 30
date="09042026"
data_type="bone_marrow_20k_sequencing_depth_cell_ranger"
```

# Generating Bootstrapped samples Copied from generating bootstrapped samples)
# In future modify the generating_bootstrapped_samples and put it in the function file. 


```{r}
# sce=sce_rds
# title=dataset
# cells_to_be_put_in_pseudobulk=cell_table %>% pull(cell_type)
# number_of_cells_in_each_celltype=cell_table %>% pull(inDeconv)
# bootstrap=1
# cell_barcode=list()

generating_a_pseudobulk_sample_seq_depth_1<-function(sce,
                                         title,
                                         cells_to_be_put_in_pseudobulk,
                                         number_of_cells_in_each_celltype, 
                                         perform_TPM_normalisation=c(TRUE,FALSE))
{ 
  # Create a S4 object with the required number of cells for each celltype
  known_sample=data.frame()
  for (i in 1:length(cells_to_be_put_in_pseudobulk))
  {
    cellcol= sce[,sce$cell==cells_to_be_put_in_pseudobulk[i]]
    if (i==1)
      known_sample= cellcol[,sample(ncol(cellcol),number_of_cells_in_each_celltype[i])]
    else
    known_sample=cbind(known_sample,cellcol[,sample(ncol(cellcol),number_of_cells_in_each_celltype[i])])
  }
  
  cell_umis <- colnames(known_sample)
  
  # Summing the counts of all the cell rowwise to get a pseudobulk sample
  samp<-as.data.frame(Matrix::rowSums(assays(known_sample)$counts))
  samp<-cbind(rownames(known_sample),samp)
  
  # Aggregating rows with common gene symbol
  samp<-aggregate(samp[,-1],by=samp[1],sum)
  
  # Getting TPM ( where TPM= (count/total count)* 1000000)
  count_sum<-sum(samp$x)
  if(perform_TPM_normalisation==TRUE)
    {
      samp[,2]<-sapply(samp[,2],function(i) (i/count_sum)*1000000)
      colnames(samp)=c("HUGO",title)
    }
  else
    {
      colnames(samp)=c("HUGO",title)
    }
  return(list(samp = samp, cell_umis = cell_umis))
}



generating_a_pseudobulk_sample_downsampled_reads<-function(sce,
                                         title,
                                         current_umis, 
                                         perform_TPM_normalisation=c(TRUE,FALSE))
{ 
  # Create a S4 object with the required number of cells for each celltype
  # known_sample=data.frame()
  # for (i in 1:length(cells_to_be_put_in_pseudobulk))
  # {
  #   cellcol= sce[,sce$cell==cells_to_be_put_in_pseudobulk[i]]
  #   if (i==1)
  #     known_sample= cellcol[,sample(ncol(cellcol),number_of_cells_in_each_celltype[i])]
  #   else
  #   known_sample=cbind(known_sample,cellcol[,sample(ncol(cellcol),number_of_cells_in_each_celltype[i])])
  # }
  # 
  # cell_umis <- colnames(known_sample)
  
  known_sample=sce[,colnames(sce) %in% current_umis]
  cell_umis_downsampled <- colnames(known_sample)
  
  # Summing the counts of all the cell rowwise to get a pseudobulk sample
  samp<-as.data.frame(Matrix::rowSums(assays(known_sample)$counts))
  samp<-cbind(rownames(known_sample),samp)
  
  # Aggregating rows with common gene symbol
  samp<-aggregate(samp[,-1],by=samp[1],sum)
  
  # Getting TPM ( where TPM= (count/total count)* 1000000)
  count_sum<-sum(samp$x)
  if(perform_TPM_normalisation==TRUE)
    {
      samp[,2]<-sapply(samp[,2],function(i) (i/count_sum)*1000000)
      colnames(samp)=c("HUGO",title)
    }
  else
    {
      colnames(samp)=c("HUGO",title)
    }
  return(list(samp = samp,cell_umis_downsampled=cell_umis_downsampled))
}
```

```{r}
# For the 100% sequencing depth sample
Generating_fixed_umis_for_pseudobulks_1_sequencing_depth <- function(dataset,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)
    {   cell_barcode=list()
        for (bootstrap in 1:bootstrap_value) 
          { 
            var_name= paste0("cell_count_", dataset)
            cell_count_table <- get(var_name)
            sce_rds <- get(dataset)
            cell_table <-data.frame("cell_type"= cell_types) %>% 
                        left_join(cell_count_table, by = c("cell_type" = "cell_type")) %>% rowwise() %>% 
                        mutate(inDeconv = as.integer(runif(1,min=10,max = Freq))) %>%  as.data.frame()
            cell_table <- cell_table %>% mutate(inDeconv_Frac = inDeconv/sum(inDeconv))
            res_loop<- generating_a_pseudobulk_sample_seq_depth_1(sce_rds, 
                                                           dataset,                              
                                                           cell_table %>% pull(cell_type), 
                                                           cell_table %>% pull(inDeconv), 
                                                           perform_TPM_normalisation=FALSE)
            
           # Adding bootstrap value to the pseudobulks generated in each run
            for_the_loop= res_loop$samp %>%
                dplyr::rename_with(~ paste0(.x,"_boot",bootstrap,".Raw"), -HUGO)
            
            cell_barcode[[paste0("boot",bootstrap,".Raw")]]=res_loop$cell_umis
        
          if(bootstrap==1)
            {
              deconv_table= for_the_loop
              actual_cell_proportions= cell_table %>% 
                             dplyr::select(cell_type,inDeconv,inDeconv_Frac) %>% 
                             dplyr::rename(!!paste0(dataset,"_boot",bootstrap,".actual_values") := "inDeconv",
                                           !!paste0(dataset,"_boot",bootstrap,".actual_frac") := "inDeconv_Frac")
                                 
            } else
            {
              deconv_table=merge(deconv_table, for_the_loop, by.x="HUGO",by.y="HUGO", all=TRUE)
              actual_cell_proportions= left_join(actual_cell_proportions, 
                                                 cell_table %>% 
                                                 dplyr::select(cell_type,inDeconv,inDeconv_Frac) %>% 
                                                 dplyr::rename(!!paste0(dataset,"_boot",bootstrap,".actual_values") := "inDeconv",
                                                               !!paste0(dataset,"_boot",bootstrap,".actual_frac") := "inDeconv_Frac"),
                                                 by=c("cell_type"="cell_type"))
           }
        
        print(paste("Bootstrap",bootstrap,"of",bootstrap_value,"of dataset",dataset,"completed"))
        }
      
  saveRDS(deconv_table,file = paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_pseudobulks_",dataset,"_",date,".RData"))
  saveRDS(actual_cell_proportions,file = paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_cell_proportions_",dataset,"_",date,".RData"))
  saveRDS(cell_barcode,file = paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_cell_umi_for_pseudobulk_",dataset,"_",date,".RData"))
}


#PBMC
Generating_fixed_umis_for_pseudobulks_1_sequencing_depth("pbmc20k_1",
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)


#Bone marrow
Generating_fixed_umis_for_pseudobulks_1_sequencing_depth("bone_marrow_20k_1",
                                            "bone_marrow_20k_sequencing_depth_cell_ranger",
                                            cell_types,
                                            bootstrap_value,
                                            date)


#test=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_cell_umi_for_pseudobulk_pbmc20k_1_",date,".RData"))

```

# Generating pseudobulks for the remaining down-sampled samples
```{r}
# 
# dataset="pbmc20k_0.00002"
# bootstrap=1
# # 
# # 
generating_bootstraped_datasets_10x_downsampled_reads <- function(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date,
                                            umi_summary_res)
{  for (dataset in dataset_included) 
    {   cell_umis_list <-readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_cell_umi_for_pseudobulk_",gsub("sequencing_depth_cell_ranger.*","",data_type),"1_",date,".RData"))
    umis_found=list()
        for (bootstrap in 1:bootstrap_value) 
          { 
            var_name= paste0("cell_count_", dataset)
            sce_rds <- get(dataset)
            
            # cell_table <-data.frame("cell_type"= cell_types) %>% 
            #             left_join(cell_count_table, by = c("cell_type" = "cell_type")) %>% rowwise() %>% 
            #             mutate(inDeconv = as.integer(runif(1,min=10,max = Freq))) %>%  as.data.frame()
            # cell_table <- cell_table %>% mutate(inDeconv_Frac = inDeconv/sum(inDeconv))
            current_umis <- unlist(cell_umis_list[[paste0("boot", bootstrap, ".Raw")]])
            loop_res<- generating_a_pseudobulk_sample_downsampled_reads(sce_rds, 
                                                           dataset, 
                                                           current_umis,
                                                           perform_TPM_normalisation=FALSE)
            
           # Adding bootstrap value to the pseudobulks generated in each run
            for_the_loop= loop_res$samp %>%
                dplyr::rename_with(~ paste0(.x,"_boot",bootstrap,".Raw"), -HUGO)
 
        
          if(dataset==dataset_included[1] && bootstrap==1)
            {
              deconv_table= for_the_loop
                                 
            } else
            {
              deconv_table=merge(deconv_table, for_the_loop, by.x="HUGO",by.y="HUGO", all=TRUE)
            }
            umis_found[[paste0("boot",bootstrap)]]=loop_res$cell_umis_downsampled
            print(paste("Bootstrap",bootstrap,"of",bootstrap_value,"of dataset",dataset,"completed"))
        }
        
    saveRDS(umis_found,file = paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_cell_umi_for_pseudobulk_",dataset,"_",date,".RData"))
    umi_summary=data.frame(n_cells   = sapply(umis_found, length)) %>% 
        rownames_to_column(var="Bootstraped_sample") %>% 
        pivot_wider(names_from = Bootstraped_sample, values_from = n_cells) %>% 
        mutate(Sample=dataset) %>% 
        select(Sample,1:30)
    
    umi_summary_res=rbind(umi_summary_res,umi_summary)
}
  decon_table_from_1=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_pseudobulks_",gsub("sequencing_depth_cell_ranger.*","",data_type),"1_",date,".RData"))
  res_decon_table= merge(deconv_table,decon_table_from_1, by.x="HUGO",by.y="HUGO", all=TRUE)
  saveRDS(res_decon_table,file = paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_pseudobulks_",date,".RData"))
  saveRDS(umi_summary_res,file = paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_umi_summary_",date,".RData"))
}

```

```{r}
## To be run with 4 cell types

# For PBMC 
pbmc_umi_summary_list=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/pbmc20k_sequencing_depth_cell_ranger_cell_umi_for_pseudobulk_pbmc20k_1_09042026.RData"))

pbmc_umi_summary=data.frame(
  n_cells   = sapply(pbmc_umi_summary_list, length)
) %>% 
  rownames_to_column(var="Bootstraped_sample") %>% 
  mutate(Bootstraped_sample=gsub(".Raw", "", Bootstraped_sample)) %>% 
  pivot_wider(names_from = Bootstraped_sample, values_from = n_cells) %>% 
  mutate(Sample="pbmc20k_1") %>% 
  select(Sample,1:30)

generating_bootstraped_datasets_10x_downsampled_reads(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date,
                                            pbmc_umi_summary)


# For Bone marrow
bone_marrow_umi_summary_list=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_cell_umi_for_pseudobulk_bone_marrow_20k_1_09042026.RData"))

bone_marrow_umi_summary=data.frame(
  n_cells   = sapply(bone_marrow_umi_summary_list, length)
) %>% 
  rownames_to_column(var="Bootstraped_sample") %>% 
  mutate(Bootstraped_sample=gsub(".Raw", "", Bootstraped_sample)) %>% 
  pivot_wider(names_from = Bootstraped_sample, values_from = n_cells) %>% 
  mutate(Sample="bone_marrow_20k_1") %>% 
  select(Sample,1:30)

generating_bootstraped_datasets_10x_downsampled_reads(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date,
                                            bone_marrow_umi_summary)


test_umi_summary=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate",data_type,"_umi_summary_",date,".RData"))
```

#For Epic (for 3 cell types)
```{r}
## PBMC
dataset_included= c("pbmc20k_0.00002","pbmc20k_0.0002","pbmc20k_0.001","pbmc20k_0.01","pbmc20k_0.2","pbmc20k_0.4","pbmc20k_0.6","pbmc20k_0.8")
cell_types= c("B cells","NK cells","T cells") 
bootstrap_value= 30
date="09042026"
data_type="pbmc20k_sequencing_depth_cell_ranger_for_epic"

# Generating the pseudobulks with 100% reads
Generating_fixed_umis_for_pseudobulks_1_sequencing_depth("pbmc20k_1",
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)

# Using the pseudobulks proportions form 100% read with down-sampled samplels
pbmc_umi_summary_list_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/pbmc20k_sequencing_depth_cell_ranger_for_epic_cell_umi_for_pseudobulk_pbmc20k_1_09042026.RData"))

pbmc_umi_summary_for_epic=data.frame(
  n_cells   = sapply(pbmc_umi_summary_list_for_epic, length)
) %>% 
  rownames_to_column(var="Bootstraped_sample") %>% 
  mutate(Bootstraped_sample=gsub(".Raw", "", Bootstraped_sample)) %>% 
  pivot_wider(names_from = Bootstraped_sample, values_from = n_cells) %>% 
  mutate(Sample="pbmc20k_1") %>% 
  select(Sample,1:30)

generating_bootstraped_datasets_10x_downsampled_reads(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date,
                                            pbmc_umi_summary_for_epic)



## Bone Marrow
dataset_included= c("bone_marrow_20k_0.00002","bone_marrow_20k_0.0002","bone_marrow_20k_0.001","bone_marrow_20k_0.01","bone_marrow_20k_0.2","bone_marrow_20k_0.4","bone_marrow_20k_0.6","bone_marrow_20k_0.8")
cell_types= c("B cells","NK cells","T cells") 
bootstrap_value= 30
date="09042026"
data_type="bone_marrow_20k_sequencing_depth_cell_ranger_for_epic"

# Generating the pseudobulks with 100% reads
Generating_fixed_umis_for_pseudobulks_1_sequencing_depth("bone_marrow_20k_1",
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date)

bone_marrow_umi_summary_list_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_for_epic_cell_umi_for_pseudobulk_bone_marrow_20k_1_09042026.RData"))

bone_marrow_umi_summary_for_epic=data.frame(
  n_cells   = sapply(bone_marrow_umi_summary_list_for_epic, length)
) %>% 
  rownames_to_column(var="Bootstraped_sample") %>% 
  mutate(Bootstraped_sample=gsub(".Raw", "", Bootstraped_sample)) %>% 
  pivot_wider(names_from = Bootstraped_sample, values_from = n_cells) %>% 
  mutate(Sample="bone_marrow_20k_1") %>% 
  select(Sample,1:30)

generating_bootstraped_datasets_10x_downsampled_reads(dataset_included,
                                            data_type,
                                            cell_types,
                                            bootstrap_value,
                                            date,
                                            bone_marrow_umi_summary_for_epic)
```

# Creating cell proportions
```{r}
# Function to create cell counts from the 100% reads for the downsampled samples
replicate_cell_count_cols_for_samples <- function(df, new_sample_names) {
   first_col <- colnames(df)[2]
  original_sample <- sub("_boot.*", "", first_col)
  
  # get the original col only one time before the loop
  original_cols <- df[, -1]
  
  for (sample in new_sample_names) {
    new_cols <- original_cols
    colnames(new_cols) <- gsub(original_sample, sample, colnames(new_cols), fixed = TRUE)
    df <- cbind(df, new_cols)
  }
  
  return(df)
}

# Saving the cell proportions for the deconv table
# PBMC
pbmc20k_1_cell_counts= readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/pbmc20k_sequencing_depth_cell_ranger_cell_proportions_pbmc20k_1_09042026.RData")) %>%
                             dplyr::rename_with(~str_replace(., "_boot$", ""), ends_with("_boot"))

pbmc_bootstraped_data_cell_counts= replicate_cell_count_cols_for_samples(pbmc20k_1_cell_counts, c("pbmc20k_0.00002","pbmc20k_0.0002","pbmc20k_0.001","pbmc20k_0.01","pbmc20k_0.2","pbmc20k_0.4","pbmc20k_0.6","pbmc20k_0.8"))

write_rds(pbmc_bootstraped_data_cell_counts, paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/pbmc20k_sequencing_depth_cell_ranger_cell_proportions_09042026.RData"))

# For Epic
pbmc20k_1_for_epic_cell_counts= readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/pbmc20k_sequencing_depth_cell_ranger_for_epic_cell_proportions_pbmc20k_1_09042026.RData")) %>%
                             dplyr::rename_with(~str_replace(., "_boot$", ""), ends_with("_boot"))

pbmc_for_epic_bootstraped_data_cell_counts= replicate_cell_count_cols_for_samples(pbmc20k_1_for_epic_cell_counts, c("pbmc20k_0.00002","pbmc20k_0.0002","pbmc20k_0.001","pbmc20k_0.01","pbmc20k_0.2","pbmc20k_0.4","pbmc20k_0.6","pbmc20k_0.8"))

write_rds(pbmc_for_epic_bootstraped_data_cell_counts, paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/pbmc20k_sequencing_depth_cell_ranger_for_epic_cell_proportions_09042026.RData"))

# Bone marrow

bone_marrow_20k_1_cell_counts= readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_cell_proportions_bone_marrow_20k_1_09042026.RData")) %>%
                             dplyr::rename_with(~str_replace(., "_boot$", ""), ends_with("_boot"))

bone_marrow_bootstraped_data_cell_counts= replicate_cell_count_cols_for_samples(bone_marrow_20k_1_cell_counts, c("bone_marrow_20k_0.00002","bone_marrow_20k_0.0002","bone_marrow_20k_0.001","bone_marrow_20k_0.01","bone_marrow_20k_0.2","bone_marrow_20k_0.4","bone_marrow_20k_0.6","bone_marrow_20k_0.8"))

write_rds(bone_marrow_bootstraped_data_cell_counts, paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_cell_proportions_09042026.RData"))

# For epic

bone_marrow_20k_1_for_epic_cell_counts= readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_for_epic_cell_proportions_bone_marrow_20k_1_09042026.RData")) %>%
                             dplyr::rename_with(~str_replace(., "_boot$", ""), ends_with("_boot"))

bone_marrow_for_epic_bootstraped_data_cell_counts= replicate_cell_count_cols_for_samples(bone_marrow_20k_1_for_epic_cell_counts, c("bone_marrow_20k_0.00002","bone_marrow_20k_0.0002","bone_marrow_20k_0.001","bone_marrow_20k_0.01","bone_marrow_20k_0.2","bone_marrow_20k_0.4","bone_marrow_20k_0.6","bone_marrow_20k_0.8"))

write_rds(bone_marrow_for_epic_bootstraped_data_cell_counts, paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_for_epic_cell_proportions_09042026.RData"))

```

# Analysis
PBMC
```{r}
date="09042026"
data_type="pbmc20k_sequencing_depth_cell_ranger"

# Pseudobulks generated from Chromium 10X
bootstraped_data=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_pseudobulks_09042026.RData"))
bootstraped_data_cell_counts=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_cell_proportions_09042026.RData"))


bootstraped_data_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_for_epic_pseudobulks_09042026.RData"))
bootstraped_data_cell_counts_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_for_epic_cell_proportions_09042026.RData"))

# Signature Matrix
#signature_reference <- readRDS("../intermediate_data/processed_sces/sequencing_depth/PBMC20k_1_cell_ranger.rds")
signature_reference <- readRDS("../intermediate_data/processed_sces/pbmc8k.rds")
analysis_type="PBMC"

# table(signature_reference$cell)
```

Bone Marrow
```{r}
data_type="bone_marrow_20k_sequencing_depth_cell_ranger"

# Pseudobulks generated from Chromium 10X
bootstraped_data=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_pseudobulks_09042026.RData"))
bootstraped_data_cell_counts=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_cell_proportions_09042026.RData")) %>%
                             dplyr::rename_with(~str_replace(., "_boot$", ""), ends_with("_boot"))
bootstraped_data_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_for_epic_pseudobulks_09042026.RData"))
bootstraped_data_cell_counts_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/",data_type,"_for_epic_cell_proportions_09042026.RData")) %>%
                             dplyr::rename_with(~str_replace(., "_boot$", ""), ends_with("_boot"))

# Signature Matrix
#signature_reference <- readRDS("../intermediate_data/processed_sces/sequencing_depth/bone_marrow_20k_1_cell_ranger.rds")
signature_reference <- readRDS("../intermediate_data/processed_sces/AML0210X.rds")

# Single the signature doesn not containbarcode and rctd expects barcode
colData(signature_reference)$Barcode <- seq_len(ncol(signature_reference))
colnames(signature_reference)=signature_reference$Barcode

signature_reference$Barcode=colnames(signature_reference)

analysis_type="bone_marrow"
```


```{r}
#date= "09042026"
#column_character_specifications_for_mcp=c("0.0002","0.0002","0.001","0.01","0.2","0.4","0.6","_1_")

column_character_specifications=c("0.00002","0.0002","0.001","0.01","0.2","0.4","0.6","0.8","_1_")
#column_character_specifications=c("0.8","_1_")
#i="0.8"

complete_sequencing_depth_benchmarking_list =list(names(column_character_specifications)) 
for (i in column_character_specifications) 
{
deconv_table=bootstraped_data %>%
             select(HUGO, contains(i)) %>% 
                 mutate(across(-1, ~ (.x / sum(.x, na.rm = TRUE)) * 1e6))

deconv_table_epic=bootstraped_data_epic %>%
                  select(HUGO,contains(i)) %>% 
                 mutate(across(-1, ~ (.x / sum(.x, na.rm = TRUE)) * 1e6))

actual_cell_proportions= bootstraped_data_cell_counts%>%
                  select(cell_type,contains(i))
actual_cell_proportions_epic= bootstraped_data_cell_counts_epic%>%
                  select(cell_type,contains(i))
    
# Deconvolution
result_df <- deconvolution_for_bootstrap_v4(deconv_table, 
                                            deconv_tabe_epic,
                                            signature_reference,
                                            actual_cell_proportions,
                                            actual_cell_proportions_epic,
                                            i)

# Wrangling_data
comp_sequencing_depth_res= result_df %>%
      pivot_longer(cols = -c(deconv_type, cell_type),
                  names_to = c("Dataset", ".value"),
                  names_pattern = "^(.*_boot\\d+)\\.(Raw|actual_values|actual_frac)$") %>% 
                  # for real data
                  # names_sep = "\\.") %>% 
     dplyr::rename("Deconvolution_fraction"="Raw") %>% 
     dplyr::filter(!deconv_type %in% c("SpatialDecon (Custom)\n cells known","SpatialDecon (TME)\n cells known","SpatialDecon (Custom)\n cells unknown")) %>% 
                    mutate(deconv_type = ifelse(deconv_type=="SpatialDecon (TME)\n cells unknown","SpatialDecon",deconv_type)) %>% 
                    mutate(Sequencing_depth=i)
  
    
    # Adding to list
    complete_sequencing_depth_benchmarking_list[[i]] = comp_sequencing_depth_res
    print(paste(i,"Sequencing depth completed"))
}

#complete_sequencing_depth_benchmarking_list$`0.00002`


#cell2loc_sequencing_depth_res <- read_csv("../intermediate_data/cell2location_res/PBMC_sequencing_depth_09042026_benchmarking_results.csv")
cell2loc_sequencing_depth_res <- read_csv("../intermediate_data/cell2location_res/aml_sequencing_depth_09042026_benchmarking_results.csv")

# Wrangling_data
sequencing_depth_res <- rbind(complete_sequencing_depth_benchmarking_list$`0.00002`,
          complete_sequencing_depth_benchmarking_list$`0.0002`,
          complete_sequencing_depth_benchmarking_list$`0.001`,
          complete_sequencing_depth_benchmarking_list$`0.01`,
          complete_sequencing_depth_benchmarking_list$`0.2`,
          complete_sequencing_depth_benchmarking_list$`0.4`,
          complete_sequencing_depth_benchmarking_list$`0.6`,
          complete_sequencing_depth_benchmarking_list$`0.8`,
          complete_sequencing_depth_benchmarking_list$`_1_`) %>% 
  mutate(Sequencing_depth= case_when(str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.00002"~"0.002% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.0002"~"0.02% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.001"~"0.1% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.01"~"1% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.2"~"20% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.4"~"40% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.6"~"60% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="0.8"~"80% reads",
                                       str_extract(Dataset, "(?<=_)[0-9.]+(?=_)")=="1"~"100% reads")) %>% 
  rbind(cell2loc_sequencing_depth_res) %>% 
                    mutate(deconv_type=factor(deconv_type, levels=c("quanTIseq",
                                                                  "EPIC",
                                                                  "MCPcounter*",
                                                                  "SPOTlight",
                                                                  "SpatialDecon",
                                                                  "RCTD",
                                                                  "cell2location"))) %>% 
    mutate(Sequencing_depth=factor(Sequencing_depth, levels=c("0.002% reads","0.02% reads","0.1% reads","1% reads","20% reads","40% reads","60% reads","80% reads","100% reads")))

date="09042026"
save(sequencing_depth_res, file = paste0("../Results/Sequencing_depth/cell_ranger/",analysis_type,"_sequencing_depth_",date,".RData"))
```

```{r}
load("../Results/Sequencing_depth/cell_ranger/PBMC_sequencing_depth_pbmc20k_21052025.RData") 
```

# Checking the depth of the pseudo-bulk sample
```{r}
sequencing_depth_of_pseudo_bulks <- tibble(
  "Sequencing_depth_prop" = rev(c("0.00002", "0.0002", "0.01", "0.2", "0.6", "_1_")),
  "Sequencing depth \n of pseudo-bulk samples" = rev(c("0.002% reads", "0.02% reads", "1% reads", "20% reads", "60% reads", "100% reads"))
)

bootstraped_data_pbmc=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/PBMC20k_sequencing_depth_cell_ranger_pseudobulks_09042026.RData"))
bootstraped_data_pbmc_cell=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/PBMC20k_sequencing_depth_cell_ranger_cell_proportions_09042026.RData"))


bootstraped_data_pbmc_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/PBMC20k_sequencing_depth_cell_ranger_for_epic_pseudobulks_09042026.RData"))
bootstraped_data_pbmc_cell_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/PBMC20k_sequencing_depth_cell_ranger_for_epic_cell_proportions_09042026.RData"))


bootstraped_data_bone_marrow=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_pseudobulks_09042026.RData"))
bootstraped_data_bone_marrow_cells=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_cell_proportions_09042026.RData"))

bootstraped_data_bone_marrow_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_for_epic_pseudobulks_09042026.RData"))
bootstraped_data_bone_marrow_cells_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_for_epic_cell_proportions_09042026.RData"))

#i="0.00002"


for (i in sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop) 
{
  # PBMC 4 cells
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean nb. of UMIs per pseudobulk sample"] <-
    bootstraped_data_pbmc %>%
      select(contains(i)) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean nb. of cells per pseudobulk sample"] <-
    bootstraped_data_pbmc_cell %>%
      select(contains(i)) %>%
      select(contains("actual_values")) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean nb. of genes per pseudobulk sample"] <-
    bootstraped_data_pbmc %>%
      select(contains(i)) %>%
      summarise(across(everything(), ~ sum(. != 0))) %>%  # count non-zero per column
      unlist() %>%                                       # convert to numeric vector
      mean() 
  
  # PBMC 3 cells
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean nb. of UMIs per pseudobulk sample (EPIC)"] <-
    bootstraped_data_pbmc_for_epic %>%
      select(contains(i)) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean nb. of cells per pseudobulk sample (EPIC)"] <-
    bootstraped_data_pbmc_cell_for_epic %>%
      select(contains(i)) %>%
      select(contains("actual_values")) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "PBMC20k \nmean nb. of genes per pseudobulk sample (EPIC)"] <-
    bootstraped_data_pbmc_for_epic %>%
      select(contains(i)) %>%
      summarise(across(everything(), ~ sum(. != 0))) %>%  # count non-zero per column
      unlist() %>%                                       # convert to numeric vector
      mean() 
  
  # Bone Marrow 4 cells
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean nb. of UMIs per pseudobulk sample"] <-
    bootstraped_data_bone_marrow %>%
      select(contains(i)) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean nb. of cells per pseudobulk sample"] <-
    bootstraped_data_bone_marrow_cells %>%
      select(contains(i)) %>%
      select(contains("actual_values")) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean nb. of genes per pseudobulk sample"] <-
    bootstraped_data_bone_marrow %>%
      select(contains(i)) %>%
      summarise(across(everything(), ~ sum(. != 0))) %>%  # count non-zero per column
      unlist() %>%                                       # convert to numeric vector
      mean() 
  
  # Bone Marrow 3 cells
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean nb. of UMIs per pseudobulk sample (EPIC)"] <-
    bootstraped_data_bone_marrow_for_epic %>%
      select(contains(i)) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean nb. of cells per pseudobulk sample (EPIC)"] <-
    bootstraped_data_bone_marrow_cells_for_epic %>%
      select(contains(i)) %>%
      select(contains("actual_values")) %>%
      colSums() %>%
      mean()
  
  sequencing_depth_of_pseudo_bulks[sequencing_depth_of_pseudo_bulks$Sequencing_depth_prop == i, "20k Bone Marrow \nmean nb. of genes per pseudobulk sample (EPIC)"] <-
    bootstraped_data_bone_marrow_for_epic %>%
      select(contains(i)) %>%
      summarise(across(everything(), ~ sum(. != 0))) %>%  # count non-zero per column
      unlist() %>%                                       # convert to numeric vector
      mean() 
}

sequencing_depth_of_pseudo_bulks=sequencing_depth_of_pseudo_bulks %>% 
  mutate("PBMC20k \nmean nb. of reads per cell"= c(53529,32396, 10984,574,11,2),
         "PBMC20k \nmean nb. of reads per pseudobulk sample"=`PBMC20k \nmean nb. of reads per cell`*`PBMC20k \nmean nb. of cells per pseudobulk sample`,
         "PBMC20k \nmean nb. of reads per pseudobulk sample (EPIC)"=`PBMC20k \nmean nb. of reads per cell`*`PBMC20k \nmean nb. of cells per pseudobulk sample (EPIC)`,
          "20k Bone Marrow \nmean nb. of reads per cell"= c(38316,23047, 7748,399,9,2),
         "20k Bone Marrow \nmean nb. of reads per pseudobulk sample"=`20k Bone Marrow \nmean nb. of reads per cell`*`20k Bone Marrow \nmean nb. of cells per pseudobulk sample`,
         "20k Bone Marrow \nmean nb. of reads per pseudobulk sample (EPIC)"=`20k Bone Marrow \nmean nb. of reads per cell`*`20k Bone Marrow \nmean nb. of cells per pseudobulk sample (EPIC)`) %>% 
  select(-c(Sequencing_depth_prop,`PBMC20k \nmean nb. of cells per pseudobulk sample`,`PBMC20k \nmean nb. of cells per pseudobulk sample (EPIC)`,
            `20k Bone Marrow \nmean nb. of cells per pseudobulk sample`,`20k Bone Marrow \nmean nb. of cells per pseudobulk sample (EPIC)`))

saveRDS(sequencing_depth_of_pseudo_bulks,"../manuscript/figure_assembly_scripts/fig_sequencing_depth/sequencing_depth_summary_numbers.RData")

library(ggplot2)
library(gridExtra)

png("../manuscript/figure_assembly_scripts/fig_sequencing_depth/sequencing_depth_summary_numbers.png", height=180, width=800,pointsize = 16)
p<-tableGrob(sequencing_depth_of_pseudo_bulks)
grid.arrange(p)
dev.off()
```


```{r}
pbmc_cell_summary=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/pbmc20k_sequencing_depth_cell_ranger_umi_summary_09042026.RData")) %>% 
                  mutate(Data="nb_cells")


test=colSums(bootstraped_data[,-1]) %>% 
     as.data.frame() %>% 
     dplyr::rename("n_count"=".") %>% 
     rownames_to_column(var="Sample_id") %>% 
     mutate(Sample=sub("_boot.*", "", Sample_id),
            Boot=sub(".*_(boot[^.]+).*", "\\1", Sample_id)) %>% 
     select(-Sample_id) %>% 
     pivot_wider(names_from = Boot, values_from = n_count) %>% 
     mutate(Data="nb_counts") %>% 
     rbind(pbmc_cell_summary) %>% 
     mutate(Sample=factor(Sample, levels = c("pbmc20k_1","pbmc20k_0.6","pbmc20k_0.4","pbmc20k_0.2","pbmc20k_0.01","pbmc20k_0.001","pbmc20k_0.0002","pbmc20k_0.00002"))) %>% 
group_by(Sample,Data)
```

# Pseudobulk summary
```{r}

bootstraped_data_pbmc_cell=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/pbmc_10x_cell_proportions_03072024.RData"))
bootstraped_data_pbmc_cell_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/pbmc_10x_for_epic_cell_proportions_13102025.RData"))



bootstraped_data_bone_marrow_cell=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/aml_10x_cell_proportions_03072024.RData"))
bootstraped_data_bone_marrow_cell_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/aml_10x_for_epic_cell_proportions_13102025.RData"))

#df <-bootstraped_data_bone_marrow_cell_for_epic
 
generating_pseudobulk_cell_composition_summary_gene_panel <- function(df, tissue, fig,nb_cells) {
  summary <- df %>%
    select(contains("actual_values")) %>%
    summarise(across(everything(), sum)) %>%
    rename_with(~ gsub(".actual_values", "", .x)) %>%
    summarise(across(everything(), sum)) %>% 
    pivot_longer(cols = 1:ncol(.), names_to = "Sample_id", values_to = "Nb_of_cells") %>%
    summarise(
      min  = min(Nb_of_cells),
      max  = max(Nb_of_cells),
      mean = mean(Nb_of_cells)
    ) %>%
    mutate(tissue = tissue, fig = fig, nb_cells=nb_cells)

  return(summary)
}

# Gene Panel 
pbmc_4_cells<- generating_pseudobulk_cell_composition_summary_gene_panel(bootstraped_data_pbmc_cell,"PBMC","Fig 2,5,S3",4) # mean= 2477.933, min= 446, max= 6713
pbmc_3_cells <- generating_pseudobulk_cell_composition_summary_gene_panel(bootstraped_data_pbmc_cell_for_epic,"PBMC","Fig 2,S3",3) # 1885.175, min= 201, max= 5169
bone_marrow_4_cells <- generating_pseudobulk_cell_composition_summary_gene_panel(bootstraped_data_bone_marrow_cell,"Bone Marrow","Fig S1,5,S3",4) #686.1778, min= 178, max=1402
bone_marrow_3_cells <- generating_pseudobulk_cell_composition_summary_gene_panel(bootstraped_data_bone_marrow_cell_for_epic,"Bone Marrow","Fig S1,S3",3) # 627.0556, min=127, max= 1410
 
# Sequencing
sequencing_depth_of_pseudo_bulks <- tibble(
  "Sequencing_depth_prop" = rev(c("0.00002", "0.0002", "0.01", "0.2", "0.6", "_1_")),
  "Sequencing depth \n of pseudo-bulk samples" = rev(c("0.002% reads", "0.02% reads", "1% reads", "20% reads", "60% reads", "100% reads"))
)


seq_pbmc_cell=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/PBMC20k_sequencing_depth_cell_ranger_cell_proportions_09042026.RData"))

seq_pbmc_cell_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/PBMC20k_sequencing_depth_cell_ranger_for_epic_cell_proportions_09042026.RData"))

seq_bone_marrow_cells=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_cell_proportions_09042026.RData"))

seq_bone_marrow_cells_for_epic=readRDS(paste0("../intermediate_data/bootstraped_pseudobulks/sequencing_depth_intermediate/bone_marrow_20k_sequencing_depth_cell_ranger_for_epic_cell_proportions_09042026.RData"))


generating_pseudobulk_cell_composition_summary_subset <- function(df, tissue, fig, substring,nb_cells) {
  summary <- df %>%
    select(contains("actual_values")) %>%
    rename_with(~ gsub(".actual_values", "", .x)) %>%
    select(contains(substring)) %>%
    summarise(across(everything(), sum)) %>% 
    pivot_longer(cols = 1:ncol(.), names_to = "Sample_id", values_to = "Nb_of_cells") %>%
    summarise(
      min  = min(Nb_of_cells),
      max  = max(Nb_of_cells),
      mean = mean(Nb_of_cells)
    ) %>%
    mutate(tissue = tissue, fig = fig,nb_cells=nb_cells)

  return(summary)
}

seq_pbmc_4 <- generating_pseudobulk_cell_composition_summary_subset(seq_pbmc_cell,"PBMC","Fig 3, S4","_1_",4)
seq_pbmc_3 <- generating_pseudobulk_cell_composition_summary_subset(seq_pbmc_cell,"PBMC","Fig 3, S4","_1_",3)
seq_bm_4 <- generating_pseudobulk_cell_composition_summary_subset(seq_pbmc_cell,"Bone marrow","Fig 3, S4","_1_",4)
seq_bm_3 <- generating_pseudobulk_cell_composition_summary_subset(seq_pbmc_cell,"Bone marrow","Fig S2","_1_",3)

spot_pbmc_cells_4 <- readRDS("../intermediate_data/bootstraped_pseudobulks/pbmc_cell_proportions_spot_size_01102025.RData") 
spot_pbmc_cells_3 <- readRDS("../intermediate_data/bootstraped_pseudobulks/pbmc_for_epic_cell_proportions_spot_size_13102025.RData") 

spot_pbmc_4_1000 <- generating_pseudobulk_cell_composition_summary_subset(spot_pbmc_cells_4,"PBMC","Fig 4",".1000_",4)
spot_pbmc_3_1000 <- generating_pseudobulk_cell_composition_summary_subset(spot_pbmc_cells_3,"PBMC","Fig 4",".1000_",3)
spot_pbmc_4_125 <- generating_pseudobulk_cell_composition_summary_subset(spot_pbmc_cells_4,"PBMC","Fig 4",".125_",4)
spot_pbmc_3_125 <- generating_pseudobulk_cell_composition_summary_subset(spot_pbmc_cells_3,"PBMC","Fig 4",".125",3)
spot_pbmc_4_4 <- generating_pseudobulk_cell_composition_summary_subset(spot_pbmc_cells_4,"PBMC","Fig 4",".4_",4)
spot_pbmc_3_4 <- generating_pseudobulk_cell_composition_summary_subset(spot_pbmc_cells_3,"PBMC","Fig 4",".4_",3)


summary_spp <- rbind(pbmc_4_cells,pbmc_3_cells,seq_pbmc_4,seq_pbmc_3,spot_pbmc_4_1000,spot_pbmc_3_1000,spot_pbmc_4_125,spot_pbmc_3_125,spot_pbmc_4_4,spot_pbmc_3_4,bone_marrow_4_cells,bone_marrow_3_cells,seq_bm_4,seq_bm_3) %>% 
  select(fig, tissue, nb_cells, min, max, mean) %>% 
  dplyr::rename("Used in Figure"=fig,
                "Tissue type"=tissue,
                "Types of cells"=nb_cells,
                "Mininum nb. of cells in a pseudobulk"=min,
                "Maximum nb. of cells in a pseudobulk"=max,
                "Mean nb. of cells in all pseudobulks"=mean)

write_csv(summary_spp, file = "Supplementary_table_4.csv")
```

```{r}
max(summary$Nb_of_cells)
```

